Weak gravitational lensing shear estimates with \textsc{metacalibration} for the \emph{Roman} High-Latitude Imaging Survey

### 2022-02-23 paper_v3.tex
v3.2

% Change log
%
% v3.0 May 2015
%    Renamed to match the new package name
%    Version number matches mnras.cls
%    A few minor tweaks to wording
% v1.0 September 2013
%    Beta testing only - never publicly released
%    First version: a simple (ish) template for creating an MNRAS paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[fleqn,usenatbib]{mnras}

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

@@

% Abstract of the paper
\begin{abstract}
We investigate the performance of the \textsc{metacalibration} shear calibration framework using image simulations for the \emph{Nancy Grace Roman} Space Telescope (\emph{Roman}) reference High-Latitude Imaging Survey (HLIS). These large-scale realistic image simulations are necessary to measure shear bias to confirm that the requirements of the instrument and the overall weak lensing survey strategy are met. The weak lensing program of the \emph{Roman} mission requires the galaxy shapes to be calibrated within about 0.03\%. To reach this goal, we can test our calibration process with various simulations and ultimately isolate the sources where these residual shear biases arise. In this work, we incorporate several new realistic processing-pipeline updates necessary to calibrate shear with a better precision relative to the previous version of the simulations and show first results for the reference HLIS using \textsc{metacalibration}. We find that in the simplified simulations where objects are drawn with the Gaussian light profile, \textsc{metacalibration} can calibrate shapes within the requirement, $m=(-0.01\pm 0.10)$\%. However, \textsc{metacalibration} as used in the most-realistic version of the simulations can only reduce the shear calibration bias to the percent level. With the single-epoch images, across all three filters we estimate the multiplicative bias $m=(-1.34\pm 0.67)$\%. This is we believe caused by the undersampled \emph{Roman} images. \textcolor{red}{In an attempt to reduce the effect of undersampling, we esimate the bias on the coadd cutouts and find that $m=(-2.33\pm 0.64)$\%.} Overall, within the current simulation volume, there is evidence that \textsc{metacalibration} as currently implemented will not calibrate shapes with zero shear bias in the \emph{Roman} HLIS and further work on the shear calibration methodology is necessary. In the future, we will need to substantially increase the simulation volume by 400 (about the same size as the survey area of 2000 deg$^{2}$) to reach the precision necessary for the HLIS.
\end{abstract}

@@

Since the formulation of a physical cosmological model, more observational evidence has shown that the contents of the universe are mostly in the dark sector (\citealt{2020A&A...633A..69H, 2018PhRvD..98d3528T, 2019PhRvL.122q1301A, 2019PASJ...71...43H, 2021A&A...645A.104A}). Within this sector, dark energy accounts for the accelerating expansion of the universe (e.g., \citealt{1998AJ....116.1009R, 1999AIPC..478..129P}) and the combination of dark energy and dark matter are responsible for the observed growth of large-scale structure (\citealt{2015RPPh...78h6901K, 2017grle.book.....D}). The growth can be studied by carefully reconstructing the paths of the light leaving distant galaxies and going through the regions where mass concentrates in the universe; this physical phenomenon is called gravitational lensing. “Weak” gravitational lensing causes a slight distortion to intrinsic galaxy shapes when observed from the Earth. By detecting millions of these gravitationally lensed galaxies and computing their ensemble distortion levels, we can explore the dark matter density and the amplitude of matter fluctuation in the universe (e.g., \citealt{2001PhR...340..291B}). Measuring these quantities eventually helps us learn about the history of the development of the large-scale structure in the universe. For that reason, weak lensing is one of the most powerful probes in current and near-future imaging surveys to constrain cosmological parameters with high precision.

Due to the subtlety in the lensing effect and its systematics-dominant nature in the weak regime, observational efforts have been challenging. Over the past decade, large international collaborations such as the Dark Energy Survey\footnote{\url{http://www.darkenergysurvey.org/}} (DES: \citealt{2005astro.ph.10346T}), the Hyper Supreme-Cam\footnote{\url{http://hsc.mtk.nao.ac.jp/ssp/}} (HSC: \citealt{2018PASJ...70S...4A}), and the Kilo-Degree Survey\footnote{\url{http://kids.strw.leidenuniv.nl/}} (KiDS: \citealt{2013ExA....35...25D}) have been successful at constraining the cosmological parameters, and the precision in calibrating the shapes of distant galaxies has reached a few percent (\citealt{2020arXiv201103408G, 2020arXiv201208567M, 2021A&A...645A.105G}). As we observe more area of the sky and develop better tools (observatories and algorithms), we detect more galaxies, and have already reached the point where statistical and systematic uncertainties are comparable. Thus, near-future Stage IV surveys (\citealt{2006astro.ph..9591A}) such as Euclid\footnote{\url{ http://sci.esa.int/euclid}} (\citealt{2011arXiv1110.3193L}), the Vera C. Rubin Observatory Legacy Survey of Space and Time\footnote{\url{ http://www.lsst.org}} (\emph{Rubin}: \citealt{2009arXiv0912.0201L, 2019ApJ...873..111I}), and \emph{Nancy Grace Roman} Space Telescope\footnote{\url{https://roman.gsfc.nasa.gov}} (\emph{Roman}: \citealt{2015arXiv150303757S}) require even better control of systematics. In order to understand the systematic uncertainties we face, we must develop realistic image simulations, apply pre-existing shape measurement to the simulated images, and determine any potential residual systematic effects and build a strategy for modifying the calibration methodology to mitigate them.

Based on our current knowledge, systematic biases for weak lensing science, both observational and astrophysical, can occur at all stages of the imaging survey (e.g., \citealt{2018ARA&A..56..393M}). Particularly, observational systematics can be due to: 
\begin{itemize}
    \item inhomogeneous observing strategy,
    \item instrumentation effects,
    \item post-processing pipelines such as image coaddition and object detection.
\end{itemize}

Conventionally, these observational systematic biases are realized through a number of simulations and validations, and these propagate into the final galaxy shapes we measure. We quantify this impact as shear calibration bias. There have been several shear calibration and bias mitigation efforts that were proposed by STEP (\citealt{2006MNRAS.368.1323H, 2007MNRAS.376...13M}) and GREAT (\citealt{2010MNRAS.405.2044B, 2013ApJS..205...12K, 2015MNRAS.450.2963M}) challenges, to meet the requirements for the current imaging surveys. In particular, one of the state-of-the-art self-calibration methods, \textsc{metacalibration} (\citealt{2017arXiv170202600H, 2017ApJ...841...24S}) has been shown to be able to substantially reduce the significance of shear bias. It has been confirmed that shear can be calibrated with \textsc{metacalibration} at the few-percent level in DES (\citealt{2018MNRAS.481.1149Z, 2020arXiv201103408G}), without accounting for galaxy blending and detection. Tackling the issue of blending/detection is one key step forward for the DES Y6 analysis and preparatory Rubin LSST work with the shear-dependent detection and calibration technique (\textsc{metadetection}; \citealt{2020ApJ...902..138S}). While the major advantage of \textsc{metacalibration} is that it can be directly applied to real galaxy images without needing to rely on an ensemble calibration from image simulations, limitations for future surveys are not well-known.

One possible limitation might lie in the effect of undersampled images in space-based surveys like \emph{Euclid} and \emph{Roman}, which will operate at its diffraction-limit. The PSF needs to be interpolated and estimated from well-sampled images to deconvolve and measure shapes with, because the \emph{Roman} PSF has a complex structure that cannot be captured in the original undersampled images. It is, therefore, necessary to build a robust strategy to reconstruct well-sampled images to estimate the PSF to allow unbiased shape measurement.
Recently, \cite{2021MNRAS.502.4048K} (hereafter K21) addressed this potential issue of the limitations of \textsc{metacalibration} on undersampled images in \emph{Euclid} image simulations. They found that for the \emph{Euclid} mission the shear estimate with \textsc{metacalibration} is biased by about 1$\%$. It is mentioned that their result could be extended to the \emph{Roman} mission due to similarities in the instruments and for \emph{Roman} they predicted the multiplicative bias was more than 1\%. They show that these effects can be mitigated using additional weighting kernels in the measurement.

@@

This paper is organized as follows. We first introduce the formalism of shear bias and image sampling relevant to the space-based surveys in Sec.~\ref{sec:background}. We then briefly discuss the simulation suite details we used and the updates we implemented for this project in Sec.~\ref{sec:sims}. In Sec.~\ref{sec:methods}, we show how coadditions of single exposure postage stamps are produced and how shape measurements are completed using \textsc{metacalibration}. We present the galaxy catalog properties and calibration results for single-band and multiband measurements in Sec.~\ref{sec:results}. Finally, in Sec.~\ref{sec:discussion}, we discuss how we will be able to constrain the shear bias better in terms of further updates in the image simulations and what we can conclude from this study.

\section{Background}
\label{sec:background}
In this section, we provide a brief introduction of image sampling defined through the Nyquist-Shannon Sampling Theorem, an overview of shear calibration bias within the context of weak lensing, and \textsc{metacalibration} formalism.

\subsection{Image Sampling}
Based on the Nyquist-Shannon Sampling Theorem which states that in order to reconstruct the unbiased continuous band-limited function without a loss of information, the sufficient sample rate is twice the bandlimiting frequency per second. In the context of image sampling, the criterion that the sample pixel spacing needs to satisfy is $p < \frac{1}{2u_{max}}$ ($u_{max}$ is the max. spatial bandlimiting frequency). Since the spatial bandlimiting frequency in an astronomical image is defined as, 
\begin{equation}
    u_{max} = \frac{1}{\lambda_{min}N_{f}}
\end{equation}
the sampling factor of the \emph{Roman} images is defined as, 
\begin{equation}
    Q = \frac{\lambda_{min}N_{f}}{p}, 
    \label{eqn:sampling}
\end{equation}
where $\lambda_{min}$ is the shortest wavelength of the incident light in each filter, $N_{f}$ is the focal ratio of the telescope ($N_{f}=7.8$ for \emph{Roman}) and $p$ is the pixel spacing of the sensor ($p=10\mu m$ for \emph{Roman}) (\citealt{2013PASP..125.1496S}). An image with sampling factor $Q=2$ is called a Nyquist-sampled image and an image with with sampling factor $Q<2$ is called an undersampled image. \textcolor{red}{An undersampled image... why is this an issue with weak lensing. What is the process of reconstructing better-sampled images.}

@@

@@

After processing the images in \textsc{metacalibration} and measuring the shapes in \texttt{ngmix}, we have 5 shape catalogs in each \textsc{metacalibration} shear direction. We can then compute the shear response from these catalogs and calculate the calibrated ensemble galaxy ellipticities. The shear response can be understood as the change in objects' ellipticities with respect to the change in gravitational shear. Mathematically, the observed ellipticities, $\vb*{e}$ = ($e_{1}$, $e_{2}$), can be approximated using a Taylor expansion around the two-component shear, $\vb*{\gamma}=0$. Higher terms here can be dropped assuming the gravitational shear is small (reduced shear). 
\begin{equation}
    \vb*{e} = \vb*{e}\rvert_{\vb*{\gamma}=0} + \frac{\partial \vb*{e}}{\partial \vb*{\gamma}}\bigg\rvert_{\vb*{\gamma}=0}\vb*{\gamma} + \ldots
\end{equation}
From this expression, the shear response matrix can be defined as, 
\begin{equation}
    \vb*{R} \equiv \frac{\partial \vb*{e}}{\partial \vb*{\gamma}}\bigg\rvert_{\vb*{\gamma}=0} = 
    \begin{pmatrix}
        \partial e_{1}/\partial \gamma_{1} & \partial e_{2}/\partial \gamma_{1} \\ 
        \partial e_{1}/\partial \gamma_{2} & \partial e_{2}/\partial \gamma_{2}
    \end{pmatrix}. 
\end{equation}
The ensemble mean of measured ellipticities $\langle\vb*{e}\rangle$ can then computed. Since we expect galaxies to be randomly oriented such that $\langle\vb*{e}\rangle\rvert_{\vb*{\gamma}=0}=0$, we have a relationship between the measured ellipticities and the input shear in the simulations, 
\begin{equation}
    \langle\vb*{e}\rangle \approx \langle \vb*{R}\vb*{\gamma} \rangle. 
\end{equation}

@@

The mean shear response is then used to get the \textsc{metacalibration}-calibrated shapes from the unsheared catalog. For individual objects, using Eq. \ref{eqn:linear}. we compare the input shear and the calibrated shapes to calculate multiplicative ($\vb*{M}$) and additive bias ($\vb*{C}$).

\section{Simulations}
\label{sec:sims}
The base of our image simulations of the \emph{Roman} Space Telescope is the \emph{Roman} simulation suite\footnote{\url{ https://github.com/matroxel/roman_imsim}} developed by T21, which renders realistic galaxy and star images on 18 Sensor-Chip Assemblies (SCAs) of a 2.5$\times$2.5 deg$^{2}$ patch of the sky following the observing strategy for the 5-year reference mission and Cycle7 instrument specifications\footnote{\url{https://roman.gsfc.nasa.gov/science/Roman_Reference_Information.html}}. We begin by creating a truth catalog using the simulated galaxy distribution from the Buzzard simulation (\citealt{2019arXiv190102401D}), a photometric galaxy catalog sampled from the CANDELS survey (e.g., \citealt{2019ApJ...877..117H}), and a Milky Way simulation (Galaxia; \citealt{2011ApJ...730....3S}) for star positions and magnitudes. Then, the following properties are applied to each galaxy:
\begin{itemize}
    \item positions (RA, Dec),
    \item flux within each \emph{Roman} filter (F184/H158/J129/Y106),
    \item intrinsic galaxy shapes and random orientations,
    \item flux ratios of de Vaucouleurs bulge, exponential disk, and star-forming knots, and
    \item artificial gravitational lensing shears.
\end{itemize} 
Within four identical realizations of the simulation, we use four sets of gravitational shears ($e_{1}$, $e_{2}$)=\{(+0.02, 0.00), (-0.02, 0.00), (0.00, +0.02), (0.00, -0.02)\}. This approach helps us to reduce shape and measurement noise when taking the difference in recovered shapes to compute the multiplicative bias (\citealt{2019A&A...621A...2P}).

@@

After we generate the object stamps across all the SCAs and pointings, we create Multi-Epoch Data Structure (MEDS) files in which each unique object dictionary contains information of all the exposures it appears in. These MEDS files are generated according to the Hierarchical Equal Area isoLatitude Pixelisation (HEALPix) of $n_{side}=512$. They contain all objects that are located in that region of the sky partitioned according HEALPixel\footnote{\url{https://healpix.jpl.nasa.gov/}} (\citealt{2005ApJ...622..759G, Zonca2019}). Once the objects are sorted in MEDS files, we pass these multiple exposures with the corresponding PSFs to \texttt{ngmix}\footnote{\url{ https://github.com/esheldon/ngmix}} to fit the galaxy shapes with the Gaussian mixture fitting method (\citealt{2014MNRAS.444L..25S}). This shape measurement process produces shape catalogs from which the shear calibration bias can be calculated with \textsc{metacalibration}.

\begin{figure}
	\includegraphics[width=\columnwidth]{psf_rotation.pdf}
% 	\vspace*{-3mm}
    \caption{The rotation of the Roman PSF produced by the \texttt{galsim.roman} module. The position angle (PA) which determines the rotation of the focal plane is $0^{\circ}$ and $30^{\circ}$ clockwise. These are drawn for SCA=1 and H158 bandpass at a native pixel scale. The rotation of the PSF is particularly important when an object has multiple exposures. As more exposures are rotated relative to one another on the sky, the average impact of the PSF will be rounder. This will translate to a substantially less-elliptical coadd PSF.}
    \label{fig:psfrot}
\end{figure}

\begin{figure*}
	\includegraphics[width=\textwidth]{lowsnr_highsnr_galaxies.pdf}
	%\vspace*{-10mm}
    \caption{The top row shows the first observation of the single epoch images and its coadded image, and the corresponding single-epoch and coadded oversampled PSF images for a galaxy with low signal-to-noise ratio (S/N=9.69 in a single-epoch image). The bottom row shows the same for a galaxy with high signal-to-noise ratio (S/N=1552.97 in a single-epoch image). The PSFs are shown in a log-scale. The single epoch PSFs are almost identical, since there is no visually different feature for modeling them in different dithers and SCAs. There are two features to note here. One is how coadding helps for low S/N objects, and the other feature is the coadded PSFs look more isotropic than the single-epoch PSF due to different numbers of exposures.}
    \label{fig:singlecoadd}
\end{figure*}

\subsection{Updates to Simulation Capabilities}
In order to accomplish the science goals, and build and test weak lensing calibration pipelines for in \emph{Roman}, we have continued to update the realism of our image simulations. We have implemented and made updates to the following parts of the simulation framework to be able to better test shear calibration using \textsc{metacalibration}. 
\begin{itemize}
    \setlength\itemsep{1em}
    \item \textbf{Saturation Cuts}:
    We have implemented a pixel saturation limit of 100,000 electrons. This value is an approximate pixel saturation level, but in the next generation of the simulations we will use a more accurate pixel saturation limit measured directly from the flight candidate detectors.

>>>>

% Change log
%
% v3.0 May 2015
%    Renamed to match the new package name
%    Version number matches mnras.clshttps://www.overleaf.com/project/605371a91cecb490b3d4a73e
%    A few minor tweaks to wording
% v1.0 September 2013
%    Beta testing only - never publicly released
%    First version: a simple (ish) template for creating an MNRAS paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[fleqn,usenatbib]{mnras}
% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

@@

% Abstract of the paper
\begin{abstract}
We investigate the performance of the \textsc{metacalibration} shear calibration framework using simulated imaging data for the \emph{Nancy Grace Roman} Space Telescope (\emph{Roman}) reference High-Latitude Imaging Survey (HLIS). 
%These large-scale realistic image simulations are necessary to measure shear bias to confirm that the requirements of the instrument and the overall weak lensing survey strategy are met. 
The weak lensing program of the \emph{Roman} mission requires the mean galaxy ellipticity to be calibrated within about 0.03\%. To reach this goal, we can test our calibration process with various simulations and ultimately isolate the sources where these residual shear biases might arise in order to improve our methods. In this work, we build on the first set of Roman HLIS simulated imaging data in \cite{2021MNRAS.501.2044T} to incorporate several new realistic processing-pipeline updates necessary to more accurately process the imaging data and calibrate the shear. We show the first results of this calibration for six deg$^{2}$ of the simulated reference HLIS using \textsc{metacalibration} and compare these results to measurements on more simple, faster Roman-like image simulations. In both cases, we neglect the impact of blending of objects. We find that in the simplified simulations, \textsc{metacalibration} can calibrate shapes to within $m=(-0.01\pm 0.10)$\%. When applied to the current most-realistic version of the simulations, the precision is much lower, with a \textsc{metacalibration} estimate of $m=(-1.34\pm 0.67)$\% for joint multi-band single-epoch measurements and $m=(-1.13\pm 0.60)$\% for multi-band coadd measurements. These results are all consistent with zero within 1--2$\sigma$, indicating we are currently limited by our simulated survey volume. Further work on testing the shear calibration methodology is necessary at higher precision to reach the levels of the \textit{Roman} requirements, in particular in the presence of blending. 
%In the future, we will need to substantially increase the simulation volume by $\sim$400 (about the same size as the survey area of 2000 deg$^{2}$) to reach the precision necessary to test the HLIS requirements. 
Current results demonstrate, however, that the \textsc{metacalibration} method can work on undersampled space-based imaging data at well-below the requirements levels of current weak lensing surveys.
\end{abstract}

@@

Since the formulation of a physical cosmological model, observational evidence has shown that the contents of the universe are mostly in the dark sector (\citealt{2020A&A...641A...8P, 2020MNRAS.498.2492G, 2020MNRAS.499.5527T, 2021A&A...645A.104A, 2022PhRvD.105b3520A, 2022arXiv220204077B}). Within this sector, dark energy accounts for the accelerating expansion of the universe (e.g., \citealt{1998AJ....116.1009R, 1999AIPC..478..129P}) and the combination of dark energy and dark matter are responsible for the observed growth of large-scale structure (\citealt{2015RPPh...78h6901K, 2017grle.book.....D}). The growth can be studied by carefully reconstructing the paths of the light leaving distant galaxies and going through the regions where mass concentrates in the universe; this physical phenomenon is called gravitational lensing. “Weak” gravitational lensing causes a slight distortion to intrinsic galaxy shapes called shear. By detecting millions of these gravitationally lensed galaxies and computing their ensemble shear, we can explore the dark matter density and the amplitude of matter fluctuation in the universe (e.g., \citealt{2001PhR...340..291B}). Measuring these quantities eventually helps us learn about the history of the development of the large-scale structure in the universe. For that reason, weak lensing is one of the most powerful probes in current and near-future imaging surveys to constrain cosmological parameters with high precision.

Due to the subtlety in the lensing effect and its systematics-dominant nature in the weak regime, observational efforts have been challenging. Over the past decade, large international collaborations such as the Dark Energy Survey\footnote{\url{http://www.darkenergysurvey.org/}} (DES: \citealt{2005astro.ph.10346T}), the Hyper Supreme-Cam\footnote{\url{http://hsc.mtk.nao.ac.jp/ssp/}} (HSC: \citealt{2018PASJ...70S...4A}), and the Kilo-Degree Survey\footnote{\url{http://kids.strw.leidenuniv.nl/}} (KiDS: \citealt{2013ExA....35...25D}) have been successful at constraining the cosmological parameters, and the precision in calibrating the shapes of distant galaxies has reached a few percent (\citealt{2020arXiv201103408G, 2020arXiv201208567M, 2021A&A...645A.105G}). As we observe more area of the sky and develop better tools (observatories and algorithms), we detect more galaxies, and have already reached the point where statistical and systematic uncertainties are comparable. Thus, near-future Stage IV surveys (\citealt{2006astro.ph..9591A}) such as Euclid\footnote{\url{ http://sci.esa.int/euclid}} (\citealt{2011arXiv1110.3193L}), the Vera C. Rubin Observatory Legacy Survey of Space and Time\footnote{\url{ http://www.lsst.org}} (\emph{Rubin}: \citealt{2009arXiv0912.0201L, 2019ApJ...873..111I}), and the \emph{Nancy Grace Roman} Space Telescope\footnote{\url{https://roman.gsfc.nasa.gov}} (\emph{Roman}: \citealt{2015arXiv150303757S}) require even better control of systematics. In order to understand the systematic uncertainties we face, we must develop realistic image simulations, apply existing shape measurement methods to the simulated images, and determine any potential residual systematic effects and build a strategy for modifying the calibration methodology to mitigate them.

Based on our current knowledge, systematic biases for weak lensing science, both observational and astrophysical, can occur at all stages of the imaging survey (e.g., \citealt{2018ARA&A..56..393M}). Particularly, observational systematics can be due to: 
\begin{itemize}
    \item inhomogeneous observing strategy,
    \item instrumentation effects,
    \item post-processing pipelines such as image coaddition and object detection.
\end{itemize} 
Conventionally, these observational systematic biases are characterized through a number of simulations and validation tests, and these propagate into uncertainties on the final mean galaxy shear we measure. We quantify this impact as shear calibration bias. There have been several shear calibration and bias mitigation efforts that were proposed by the STEP (\citealt{2006MNRAS.368.1323H, 2007MNRAS.376...13M}) and GREAT (\citealt{2010MNRAS.405.2044B, 2013ApJS..205...12K, 2015MNRAS.450.2963M}) challenges, to meet the requirements for the current imaging surveys. In particular, one of the state-of-the-art self-calibration methods, \textsc{metacalibration} (\citealt{2017arXiv170202600H, 2017ApJ...841...24S}) has been shown to be able to substantially reduce the significance of shear bias. It has been confirmed that shear can be calibrated with \textsc{metacalibration} at the few-percent level in DES (\citealt{2018MNRAS.481.1149Z, 2020arXiv201103408G}), without accounting for galaxy blending and detection. Tackling the issue of blending/detection is one key step forward for the DES Y6 analysis and preparatory Rubin LSST work with the shear-dependent detection and calibration technique \textsc{metadetection} (\citealt{2020ApJ...902..138S}). While the major advantage of \textsc{metacalibration} or \textsc{metadetection} is that they can be directly applied to real galaxy images without needing to rely on an ensemble calibration from image simulations, limitations for future surveys are not well-known.

One possible limitation might lie in the effect of undersampled images in space-based surveys like \emph{Euclid} and \emph{Roman}, which will operate at its diffraction-limit. The PSF needs to be interpolated and estimated from well-sampled images to accurately deconvolve and measure shapes with, because the \emph{Roman} PSF has a complex structure that cannot be captured in the original undersampled images. It is, therefore, necessary to build a robust strategy to reconstruct well-sampled images to estimate the PSF to allow unbiased shape measurement.
Recently, \cite{2021MNRAS.502.4048K} (hereafter K21) addressed this potential issue of the limitations of \textsc{metacalibration} on undersampled images in \emph{Euclid} image simulations. They found that for the \emph{Euclid} mission the shear estimate with \textsc{metacalibration} is biased by about 1$\%$. It is mentioned that their result could be extended to the \emph{Roman} mission due to similarities in the instruments and for \emph{Roman} they predicted the multiplicative bias was more than 1\%. They show that these effects can be mitigated using additional weighting kernels in the measurement.

@@

This paper is organized as follows. We first introduce the formalism of shear bias and image sampling relevant to the space-based surveys in Sec.~\ref{sec:background}. We then briefly discuss the simulation suite details we used and the updates we implemented for this project in Sec.~\ref{sec:sims}. In Sec.~\ref{sec:methods}, we show how coadditions of single exposure postage stamps are produced and how shape measurements are completed using \textsc{metacalibration}. We present the galaxy catalog properties and calibration results for single-band and multi-band measurements in Sec.~\ref{sec:results}. Finally, in Sec.~\ref{sec:discussion}, we discuss how we will be able to constrain the shear bias better in terms of further updates in the image simulations and what we can conclude from this study.

\section{Background}
\label{sec:background}
In this section, we provide a brief introduction of image sampling defined through the Nyquist-Shannon Sampling Theorem, an overview of shear calibration bias within the context of weak lensing, and the \textsc{metacalibration} formalism.

\subsection{Image Sampling}
Based on the Nyquist-Shannon Sampling Theorem which states that in order to reconstruct an unbiased continuous band-limited function without a loss of information, the sufficient sample rate is twice the bandlimiting frequency per second. In the context of image sampling, the criterion that the sample pixel spacing needs to satisfy is $p < \frac{1}{2u_{max}}$ ($u_{max}$ is the max. spatial bandlimiting frequency). Since the spatial bandlimiting frequency in an astronomical image is defined as, 
\begin{equation}
    u_{max} = \frac{1}{\lambda_{min}N_{f}}
\end{equation}
the sampling factor of the \emph{Roman} images is defined as, 
\begin{equation}
    Q = \frac{\lambda_{min}N_{f}}{p}, 
    \label{eqn:sampling}
\end{equation}
where $\lambda_{min}$ is the shortest wavelength of the incident light in each filter, $N_{f}$ is the focal ratio of the telescope ($N_{f}=7.8$ for \emph{Roman}) and $p$ is the pixel spacing of the sensor ($p=10\mu m$ for \emph{Roman}) (\citealt{2013PASP..125.1496S}). An image with sampling factor $Q=2$ is considered Nyquist-sampled and an image with with sampling factor $Q<2$ is undersampled.

% \textcolor{red}{Undersampled images qualitatively have less flux distribution information of objects than the Nyquist-sampled images. This particularly affects weak lensing experiments, because it involves measuring galaxy shapes. Shape measurement can bias the experiments when the galaxy images are deconvolved with undersampled PSF which does not capture the true flux spread among pixels. It is therefore preferred to conduct weak lensing experiments with at least Nyquist-sampled images. }

@@

\begin{figure}
	\includegraphics[width=\columnwidth]{metacal_bias_shear.pdf}
    %\vspace*{-5mm}
    \caption{We show the recovered multiplicative bias on the shear as a function of applied shear for the choice of the applied shear in our simulations. The presented multiplicative bias is the average of $m_{1}$ and $m_{2}$. At first order, the \textsc{metacalibration} algorithm only works for small input shears. The next order effect is seen as a quadratic dependence of the multiplicative shear bias on on the applied shear (\citealt{2017ApJ...841...24S}). A larger input shear increases the signal-to-noise of bias estimates, but to keep the non-linear effect small compared to requirements, we limit input shears to $\pm 0.02$. }
    \label{fig:metacal_shear_linear}
\end{figure}

@@

After processing the images in \textsc{metacalibration} and measuring the shapes in \texttt{ngmix}, we have 5 shape catalogs in each \textsc{metacalibration} shear direction. We can then compute the shear response from these catalogs and calculate the calibrated ensemble galaxy ellipticities. The shear response can be understood as the change in objects' ellipticities with respect to the change in gravitational shear. Mathematically, the observed ellipticities, $\vb*{e}$ = ($e_{1}$, $e_{2}$), can be approximated using a Taylor expansion around the two-component shear, $\vb*{\gamma}=0$. Higher terms here can be dropped assuming the gravitational shear is small (reduced shear). 
\begin{equation}
    \vb*{e} = \vb*{e}\rvert_{\vb*{\gamma}=0} + \frac{\partial \vb*{e}}{\partial \vb*{\gamma}}\bigg\rvert_{\vb*{\gamma}=0}\vb*{\gamma} + \ldots
\end{equation}
From this expression, the shear response matrix can be defined as, 
\begin{equation}
    \vb*{R} \equiv \frac{\partial \vb*{e}}{\partial \vb*{\gamma}}\bigg\rvert_{\vb*{\gamma}=0} = 
    \begin{pmatrix}
        \partial e_{1}/\partial \gamma_{1} & \partial e_{2}/\partial \gamma_{1} \\ 
        \partial e_{1}/\partial \gamma_{2} & \partial e_{2}/\partial \gamma_{2}
    \end{pmatrix}. 
\end{equation}
The ensemble mean of measured ellipticities $\langle\vb*{e}\rangle$ can then computed. Since we expect galaxies to be randomly oriented such that $\langle\vb*{e}\rangle\rvert_{\vb*{\gamma}=0}=0$, we have a relationship between the measured ellipticities and the input shear in the simulations, 
\begin{equation}
    \langle\vb*{e}\rangle \approx \langle \vb*{R}\rangle\langle\vb*{\gamma} \rangle. 
\end{equation}

@@

The mean shear response is then used to get the \textsc{metacalibration}-calibrated shapes from the unsheared catalog. For individual objects, using Eq. \ref{eqn:linear}. we compare the input shear and the calibrated shapes to calculate multiplicative ($\vb*{m}$) and additive bias ($\vb*{c}$).

\section{Simulations}
\label{sec:sims}
The base of our image simulations of the \emph{Roman} Space Telescope is the \emph{Roman} simulation suite\footnote{\url{ https://github.com/matroxel/roman_imsim}} developed by T21, which renders realistic galaxy and star images on 18 Sensor-Chip Assemblies (SCAs) of a 2.5$\times$2.5 deg$^{2}$ patch of the sky following the observing strategy for the 5-year reference mission and Cycle7 instrument specifications.\footnote{\url{https://roman.gsfc.nasa.gov/science/Roman_Reference_Information.html}} We begin by creating a truth catalog using the simulated galaxy distribution from the Buzzard simulation (\citealt{2019arXiv190102401D}), a photometric galaxy catalog sampled from the CANDELS survey (e.g., \citealt{2019ApJ...877..117H}), and a Milky Way simulation (Galaxia; \citealt{2011ApJ...730....3S}) for star positions and magnitudes. Then, the following properties are applied to each galaxy:
\begin{itemize}
    \item positions (RA, Dec),
    \item flux within each \emph{Roman} filter (F184/H158/J129/Y106),
    \item intrinsic galaxy shapes and random orientations,
    \item flux ratios of de Vaucouleurs bulge, exponential disk, and star-forming knots, and
    \item artificial gravitational lensing shears.
\end{itemize} 
Within four identical realizations of the simulation, we use four sets of gravitational shears ($e_{1}$, $e_{2}$)=\{(+0.02, 0.00), (-0.02, 0.00), (0.00, +0.02), (0.00, -0.02)\}. This approach helps us to reduce shape and measurement noise when taking the difference in recovered shapes to compute the multiplicative bias (\citealt{2019A&A...621A...2P}).

@@

After we generate the object stamps across all the SCAs and pointings, we create Multi-Epoch Data Structure (MEDS) files in which each unique object dictionary contains information of all the exposures it appears in. These MEDS files are partitioned according to the Hierarchical Equal Area isoLatitude Pixelisation (HEALPix) of $n_{side}=512$. They contain all objects that are located in that region of the sky partitioned according HEALPixel\footnote{\url{https://healpix.jpl.nasa.gov/}} (\citealt{2005ApJ...622..759G, Zonca2019}).

Once the objects are sorted in MEDS files, we pass these multiple exposures with the corresponding PSFs to \texttt{ngmix}\footnote{\url{ https://github.com/esheldon/ngmix}} to fit the galaxy shapes with the Gaussian mixture fitting method (\citealt{2014MNRAS.444L..25S}). This shape measurement process produces shape catalogs from which the shear calibration bias can be calculated with \textsc{metacalibration}.

\begin{figure}
	\includegraphics[width=\columnwidth]{psf_rotation.pdf}
% 	\vspace*{-3mm}
    \caption{The rotation of the Roman PSF produced by the \texttt{galsim.roman} module. The position angle (PA) which determines the rotation of the focal plane is $0^{\circ}$ (\textbf{left}) and $30^{\circ}$ (\textbf{right}) clockwise. These are drawn for SCA=1 and H158 bandpass at a native pixel scale. The rotation of the PSF is particularly important when an object has multiple exposures. As more exposures are rotated relative to one another on the sky, the average impact of the PSF will be rounder. This will translate to a substantially less-elliptical coadd PSF.}
    \label{fig:psfrot}
\end{figure}

\begin{figure*}
	\includegraphics[width=\textwidth]{lowsnr_highsnr_galaxies_v2.pdf}
	%\vspace*{-10mm}
    \caption{\textbf{Top Row}: the first observation of the single-epoch images and its coadded image, and the corresponding single-epoch and coadded oversampled PSF images for a galaxy with low signal-to-noise ratio (S/N=20.16 in a single-epoch image). \textbf{Bottom Row}: the same as above for a galaxy with high signal-to-noise ratio (S/N=390.47 in a single-epoch image). H158 bandpass was used to represent images here. The PSFs are shown in a log-scale. The single-epoch PSFs are almost identical, since there is no visually different feature for modeling them in different dithers and SCAs. There are two features to note here. One is how coadding helps for low S/N objects, and the other feature is the coadded PSFs look more isotropic than the single-epoch PSF due to different numbers of exposures.}
    \label{fig:singlecoadd}
\end{figure*}

\subsection{Updates to Simulation Capabilities}
In order to accomplish the science goals, and build and test weak lensing calibration pipelines for \emph{Roman}, we have continued to update the realism of our image simulations. We have implemented and made updates to the following parts of the simulation framework to be able to better test shear calibration using \textsc{metacalibration}. 
\begin{itemize}
    \setlength\itemsep{1em}
    \item \textbf{Saturation Cuts}:
    We have implemented a pixel saturation limit of 100,000 electrons. This value is an approximate pixel saturation level, but in the next generation of the simulations we will use a more accurate pixel saturation limit measured directly from the flight candidate detectors.
